Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex
ssssshit

\section{Example Properties and their Monitors\pages{1.5}}\label{sec:compprop}

This section introduces a trace model (\Cref{subsec:propdefs}) and defines some example properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
These properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
\subsection{Specification Trace Model}\label{subsec:propdefs}

@@

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre--event, a control--tag, and a security--tag.
The purpose of the model is to define properties, such as a stricter variant of cryptographic constant time.
To this end, security--tags indicate wether an event contains sensitive information ($\lock$) or not ($\unlock$), while control--tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre--events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Likewise, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone allow already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
Typical operations with data--dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture dependent, but division is an operation that serves as a classic example for a data--dependent timing instruction.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
\subsection{Monitors\pages{1.5}}\label{sec:monitors}
This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor--level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter--equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor--satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

>>>>

\section{Security Properties: Formalisation, Enforcement and Composition\pages{1.5}}\label{sec:compprop}

This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
\MPin{
	missing 3.2 ref? added below?
}
Then, this section introduces monitors for each of the properties presented earlier (\Cref{??}).

\subsection{Specification Trace Model}\label{subsec:propdefs}

@@

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
The purpose of the model is to define key security properties of interest, such as \MP{ms?} a stricter variant of cryptographic constant time.
To this end, security-tags indicate wether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state wether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}. \MP{cits?}
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone allow\MP{allow who? allow needs an object, this is improper english} already a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data--dependent must not have a secret as an operand.
Typical operations with data--dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent\MPin{hyphen needed}, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
\subsection{Monitors\pages{1.5}}\label{sec:monitors}

The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.\MP{not true. safety props are enforced by monitors}
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor--level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter--equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor--satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.
